Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-10-10 manuscript2.RMD
updates

**Conclusion:** Glucose supplementation before and after resistance training did not augment ribosomal accumulation during two weeks of heavy-load resistance training in moderately trained young adults.

@@

The purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT-sessions over a period of 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions. Secondly, we aimed to describe the association between changes in total RNA abundance and UBF in human skeletal muscle.

@@

The five-session-RT intervention led to increases in all measured proteins in both glucose and placebo (Figure 3A). RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).

Both baseline and trained state data of total RNA showed a linear relationship with UBF, where $\sim$ `r corr_incr`% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = `r corrp`; Figure 3C).

@@

The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA, and rDNA transcription initiation-associated proteins, compared to RT with placebo. Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume. Towards the end of the intervention, RT with glucose led to less reduction in strength compared to placebo, which may point to an accumulated effect of the glucose condition. As by design the levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with the glucose condition compared to placebo, and there were no differences in the daily macro nutrient intake between conditions on consecutive days, thus enabling an effective design to study the biological and functional effects of the glucose condition. These findings suggest that the effects of high vs. low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*, with a design as the present study. Notably, the present study was specifically designed to investigate differences in pre to post changes within participants and is as such also quite limited to this perspective. Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.

To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few. *In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011] and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015]. These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015]. In addition, Hillier et al. found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000]. Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo would elicit this effect. Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs. placebo condition would stimulate energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) and Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs. placebo conditions. Resistance training, irrespective of condition, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Due to the exploratory nature of our hypotheses, this was hardly a very surprising result. Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important in supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.

In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data. UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016]. Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. Hence, it seems quite reasonable to observe similar changes in these three proteins. The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis. While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022]. As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].

@@

@@

In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of heavy resistance exercise training, in moderately trained young adults. Neither does it acutely enhance muscular performance during the exercise period or after five sessions, nor recovery within 23 hours of the last session. The observations in total RNA, rRNA and rDNA transcription initiation-associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training. If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher-resolution results.

>>>>

**Conclusion:** Glucose ingestion before and after resistance training did not augment ribosomal RNA accumulation during two weeks of heavy-load resistance training in moderately trained young adults.

@@

Therefore, the purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT-sessions over a period of 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions. Secondly, we aimed to describe the association between changes in total RNA abundance and UBF in human skeletal muscle.

@@

The five-session-RT intervention led to increases in all measured proteins in both glucose and placebo (Figure 3A). RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A). Both baseline and trained state data of total RNA showed a linear relationship with UBF, where $\sim$ `r corr_incr`% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = `r corrp`; Figure 3C).

@@

The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA, and rDNA transcription initiation-associated proteins, compared to RT with placebo. Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume. Towards the end of the intervention, RT with glucose led to less reduction in strength compared to placebo, which may point to an accumulated effect of the glucose condition. As by design the levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with the glucose condition compared to placebo, and there were no differences in the daily macro nutrient intake between conditions on consecutive days, thus enabling an effective design to study the biological and functional effects of the glucose condition. These findings suggest that the effects of high vs. low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*, with a design like the present study.

To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle, making our hypotheses exploratory and comparisons few. *In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011] and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015]. These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015]. In addition, Hillier et al. found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000]. Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo would elicit this effect. Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs. placebo condition would stimulate energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) and Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs. placebo conditions. Resistance training, irrespective of condition, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Due to the exploratory nature of our hypotheses, this was hardly a very surprising result. Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.

In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data as there was no difference between the glucose and placebo conditions. UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016]. Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. Hence, it seems quite reasonable to observe similar changes in these three proteins. The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis. While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022]. As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].

@@

*Limitations and strengths*
Notably, the present study was specifically designed to investigate differences in pre to post changes within participants and is as such also quite limited to this perspective. Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.

@@

In conclusion, ingestion of glucose immediately before and after five heavy-load resistance training sessions conducted over a period of 12 days did not augment accumulation of ribosomal RNA, in moderately trained young adults compared to ingestion of placebo. Glucose ingestion did not affect muscular performance over the course of the study, nor did it affect muscular performance measured 30 minutes, 2 hours or 23 hours after the last session. Glucose ingestion was associated with less reduction in muscular performance 23 hours after the fifth training session, thus we cannot rule out a possible accumulated effect of ingestion glucose compared to placebo. The relationship between baseline and trained state data of total RNA and UBF levels do support the key role of UBF in ribosomal biogenesis regulation in human skeletal muscle following resistance training. Future investigations should aim for a greater sample size over a longer time period, to investigate the potential accumulated effects of glucose ingestion, as well as analyses of the intramuscular glycogen storage to provide higher-resolution results.
